---
title: "Dataset 1 Analysis"
author: "Prathishta"
date: "2024-10-17"
description: "Valentine's day Expense"
---

# Analysis of Valentine's day spending.

## 1.Setting up R Packages

loading the required packages:

```{r}
library(tidyverse)
library(mosaic)
library(skimr)
library(ggformula)
library(ggprism)
```

```{r}
# https://stackoverflow.com/questions/74491138/ggplot-custom-fonts-not-working-in-quarto

# Chunk options
knitr::opts_chunk$set(
  fig.width = 7,
  fig.asp = 0.618, # Golden Ratio
  # out.width = "80%",
  fig.align = "center"
)
### Ggplot Theme
### https://rpubs.com/mclaire19/ggplot2-custom-themes

theme_custom <- function() {
  font <- "Roboto Condensed" # assign font family up front

  theme_classic(base_size = 14) %+replace% # replace elements we want to change

    theme(
      panel.grid.minor = element_blank(), # strip minor gridlines
      text = element_text(family = font),
      # text elements
      plot.title = element_text( # title
        family = font, # set font family
        size = 16, # set font size
        face = "bold", # bold typeface
        hjust = 0, # left align
        # vjust = 2                #raise slightly
        margin = margin(0, 0, 10, 0)
      ),
      plot.subtitle = element_text( # subtitle
        family = font, # font family
        size = 14, # font size
        hjust = 0,
        margin = margin(2, 0, 5, 0)
      ),
      plot.caption = element_text( # caption
        family = font, # font family
        size = 8, # font size
        hjust = 1
      ), # right align

      axis.title = element_text( # axis titles
        family = font, # font family
        size = 10 # font size
      ),
      axis.text = element_text( # axis text
        family = font, # axis family
        size = 8
      ) # font size
    )
}

# Set graph theme
theme_set(new = theme_custom())
#
```

```{r}
#Reading data
gifts_age <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-02-13/gifts_age.csv")
glimpse(gifts_age)
```

## 2.Introduction

This is an analysis of Valentine's Day spending across various gift categories, broken down by different age groups. The aim is to explore patterns in spending for gifts such as candy, flowers, jewelry, and greeting cards, and how they differ by age.

## 3.Data Dictionary

-   **Quantitative Variables:**

    -   Age: Age groups of individuals (18-24, 25-34, 35-44, 45-54, 55-64, 65+)

    -   Candy: Amount spent on candy

    -   Flowers: Amount spent on flowers

    -   Jewellery: Amount spent on jewelry

    -   GiftCards: Amount spent on gift cards

    -   Clothings: Amount spent on clothing

    -   Greeting Cards: Amount spent on greeting cards

    -   EveningOuts: Amount spent on an evening out

    -   SpendingCelebrating: Total amount spent on celebrating Valentine’s Day

-   **Qualitative Variables:**

    -   None

    ## 4.Target and Predictor Variable

    -   Target Variable: Spending amounts across different categories

    -   Predictor Variable: Age group of the participants

## 5. Research Experiment

The dataset was likely collected through surveys or sales reports that categorized purchases based on age groups.

## 6.Research Questions

-   How does spending on various gift categories change across different age groups?

-   Which age group spends the most on specific categories like candy, jewelry, or evening outings?

-   Are there notable differences in spending trends for items typically associated with Valentine's Day (e.g., flowers, greeting cards)?

### 7.Data Preprocessing

```{r}
gifts_age_long <- gifts_age %>%
  pivot_longer(cols = -Age, names_to = "Item", values_to = "AmountSpent") %>%
  mutate(Age = factor(Age, levels = c("18-24", "25-34", "35-44", "45-54", "55-64", "65+")),
         Item = factor(Item, levels = c("Candy", "Clothing", "EveningOut", "Flowers", 
                                        "GiftCards", "GreetingCards", "Jewelry", "SpendingCelebrating")))

```

```{r}
ggplot(gifts_age_long, aes(x = Age, y = AmountSpent, color = Item, shape = Item)) +
  geom_line(aes(group = Item), size = 1) +
  geom_point(size = 3) +
  scale_shape_manual(values = c(0, 1, 2, 3, 4, 5, 6, 7)) +  # Different shapes for each item
  labs(title = "Valentine's Day Spending Across Age Groups",
       subtitle = "Amounts spent per gift category",
       x = "Age Group in Years",
       y = "Amount Spent (in $)",
       caption = "Data Source: Tidy Tuesday 13-Feb-2024") +
  theme_prism(axis_text_angle = 45) +  # Customizing axis with ggprism
  theme(legend.title = element_text(size = 12), 
        legend.text = element_text(size = 10),
        plot.title = element_text(face = "bold", size = 16),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12)) +
  scale_color_prism()  # Using ggprism color palette

```

## 8.Inference and Storytelling

-   Candy and Flowers: The 25-34 age group tends to spend the most on candy and flowers, possibly as they are within the prime age of gift-giving.

-   Jewelry: Spending on jewelry is highest in the 35-44 age group, which may be associated with more stable financial standing.

-   Greeting Cards and Gift Cards: Greeting cards and gift cards are more evenly distributed across age groups, with only slight peaks in specific age ranges.

-   Evening Out: Spending on evenings out peaks in the 25-34 and 35-44 age groups, showing a preference for experiences over material gifts during these years.

## 9. Conclusion

This analysis sheds light on the different spending patterns for Valentine's Day gifts across age groups. While younger individuals (18-24) may spend more on smaller, less expensive items like candy, older age groups tend to invest in higher-value gifts like jewelry or an evening out. Understanding these trends can help marketers and retailers target their promotions more effectively during the Valentine's Day season.
